---
title: "model-calibration"
author: "Alexander Keth"
date: '`r Sys.Date()`'
output: 
  pdf_document:
    toc: true
    number_sections: true
    fig_width: 25
    fig_height: 6
    fig_caption: true
classoption: landscape
geometry: margin = 0.5cm
vignette: >
  %\VignetteIndexEntry{model-calibration}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE}
library("atlantistools")
library("ggplot2")
library("gridExtra")

# NOTE: This vigentte is optimised for longer simulation runs. Therefore the output is not as pleasant
# due to the fact that the dummy setas file have a running time of 3 years.

# This section is used to read in the data created with preprocess(). 
# You can do this however you like. Either make sure the data is already
# present in your workspace as object "result" or load the data from *.rda with load().
# Unfortunately it is nearly impossible to demonstarte this with the package itself because the data is 
# already loaded as soon as you load the package with "library()". Therefore I have to 
# assign the data by hand otherwise the vignette will fail.
# Please comment out/edit the next two lines in your personal version of this *.Rmd file and add your
# personal loading technique.
result <- preprocess_setas
model_path <- system.file("extdata", "setas-model-new-becdev", package = "atlantistools")

# In addition please change the names of the following dataframes.
bench_bio <- "setas-bench.csv"
bench_rec <- "setas-ssb-rec.csv"

# And adjust the ploting sizes (here, at the top and throughout this document) to your likings.
fig_height2 <- 11

# You should be able to build the vignette either by clicking on "Knit-PDF" or with
# rmarkdown::render("model-calibration.Rmd")
```

\newpage

# Whole system plots!
## Overall biomass
```{r, fig.width = 14, fig.height = 8}
plot_ws(result$biomass)
```

\newpage

## Biomass timeseries
```{r}
plot <- plot_ts(result$biomass)
update_labels(p = plot, labels = list(y = "Biomass [t]"))
```

\newpage

## Number timeseries
```{r, fig.height = fig_height2}
plot <- plot_ts(result$nums)
update_labels(p = plot, labels = list(y = "Numbers"))
```

\newpage

## SSB & Recruitment

```{r}
ex <- read.csv(file.path(model_path, bench_rec), stringsAsFactors = FALSE)
plot_rec(result$ssb_rec, ex_data = ex)
```

\newpage

## Biomass benchmark
```{r}
ex <- read.csv(file.path(model_path, bench_bio), stringsAsFactors = FALSE)
plot_bench(result$biomass, ex_data = ex)
```

\newpage

## Physics
```{r, fig.width = 30, fig.height = 12}
plot_physics(result$physics)
```

\newpage

## Fluxes
```{r, fig.width = 30, fig.height = 12}
plot_flux(result$flux)
```

\newpage

# Calibration plots
## Structural nitrogen
```{r, fig.height = fig_height2}
plot_calibrate(result$structn_age)
```

\newpage

## Reserve nitrogen
```{r, fig.height = fig_height2}
plot_calibrate(result$resn_age)
```

\newpage

## Biomass per ageclass
```{r, fig.height = fig_height2}
plot_calibrate(result$biomass_age)
```

\newpage

## Numbers
```{r, fig.height = fig_height2}
plot_calibrate(result$nums_age)
```


\newpage

## Biomass
```{r}
plot_calibrate(result$biomass)
```

\newpage

# Distribution plots
## Numbers @ age
```{r, fig.height = fig_height2}
plot_struct(result$nums_age)
```

\newpage

## Biomass @ age
```{r, fig.height = fig_height2}
plot_struct(result$biomass_age)
```

\newpage

# Diet Plots

```{r, fig.width = 14, fig.height = 9}
# diet_data <- combine_ages(dir = model_path, data = result$diet_specmort, col = "pred", prm_biol = "VMPA_setas_biol_fishing_New.prm")
marrangeGrob(plot_diet(result$diet_specmort, wrap_col = "agecl", combine_thresh = 15), nrow = 1, ncol = 1)
```
