---
title: "model-calibration"
author: "Alexander Keth"
date: '`r Sys.Date()`'
output: pdf_document
toc: true
number_sections: true
classoption: landscape
geometry: margin = 0.5cm
---

```{r, echo = FALSE}
library("atlantistools")
library("ggplot2")

# This section is used to read in the data! You can do this however you like.
# The data should be present either in your workspace as object "result" or you
# can read the data in at this stage!
# I demonstrate this with the stetas model files which come shipped with the package!
# Solution1: Preprocess the data with "preprocess()".
d <- system.file("extdata", "setas-model-new-becdev", package = "atlantistools")
result <-   preprocess(dir = d,
                       nc_gen = "outputSETAS.nc",
                       nc_prod = "outputSETASPROD.nc",
                       dietcheck = "outputSETASDietCheck.txt",
                       yoy = "outputSETASYOY.txt",
                       ssb = "outputSETASSSB.txt",
                       prm_biol = "VMPA_setas_biol_fishing_New.prm",
                       prm_run = "VMPA_setas_run_fishing_F_New.prm",
                       bps = load_bps(dir = d, fgs = "SETasGroups.csv", 
                                      init = "init_vmpa_setas_25032013.nc"),
                       fgs = "SETasGroups.csv",
                       select_groups = c("Planktiv_S_Fish", "Pisciv_S_Fish", 
                                         "Cephalopod", "Diatom", "Zoo"),
                       bboxes = get_boundary(boxinfo = load_box(dir = d, 
                                                                bgm = "VMPA_setas.bgm")),
                       check_acronyms = TRUE,
                       modelstart = "1991-01-01",
                       out = "preprocess.Rda",
                       report = FALSE,
                       save_to_disc = FALSE)

# Solution2: Load in already preprocessed data! This can be done with load():
# All you have to to is comment in the next two lines and change the model_path object
# and the name of the *.Rda file.
# model_path <- file.path("C:", "ATLANTIS_Stuff", "Baseruns", "0961_v.11.0.0_ATLANTIS_NS")
# load(file.path(model_path, "ns_preprocess.Rda"))

# At this point the data is read in as object "result".

# Unfortunately it is nearly impossible to demonstarte this with the package itself because the data is 
# already loaded as soon as you load the package with "library()". Therefore I have to 
# assign the data by hand otherwise the vignette will fail.
# Please comment out the next line in your personal version of this *.Rmd file.
result <- preprocess_setas

# All you have to do now is change the names of the following dataframes.
bench_bio <- "setas-bench.csv"
bench_rec <- "setas-ssb-rec.csv"

# And adjust the ploting sizes (here and throughout this document) to your likings.
dw1 <- 25
dh1 <- 15
dh2 <- 11

# You should be able to build the vignette either by clicking on "Knit-PDF" or with
# rmarkdown::render("model-calibration.Rmd")
```

\newpage

## Whole system plots!
```{r, fig.width = 14, fig.height = 8}
plot_ws(result$biomass)
```

\newpage

## Biomass timeseries
```{r, fig.width = dw1, fig.height = dh1}
plot <- plot_ts(result$biomass)
update_labels(p = plot, labels = list(y = "Biomass [t]"))
```

\newpage

## Number timeseries
```{r, fig.width = dw1, fig.height = dh2}
plot <- plot_ts(result$nums)
update_labels(p = plot, labels = list(y = "Numbers"))
```

\newpage

## SSB & Recruitment
```{r, fig.width = dw1, fig.height = dh1}
ex <- read.csv(file.path(model_path, bench_rec), stringsAsFactors = F)
plot_rec(result$ssb_rec, ex_data = ex)
```

\newpage

## Biomass benchmark
```{r, fig.width = dw1, fig.height = dh1}
ex <- read.csv(file.path(model_path, bench_bio), stringsAsFactors = F)
plot_bench(result$biomass, ex_data = ex)

```

\newpage

## Physics
```{r, fig.width = 30, fig.height = 12}
plot_physics(result$physics)
```

\newpage

## Fluxes
```{r, fig.width = 30, fig.height = 12}
plot_flux(result$flux)
```

\newpage

## Calibration plots
### Structural nitrogen
```{r, fig.width = dw1, fig.height = dh2}
plot_calibrate(result$structn_age)
```

\newpage

### Reserve nitrogen
```{r, fig.width = dw1, fig.height = dh2}
plot_calibrate(result$resn_age)
```

\newpage

### Biomass per ageclass
```{r, fig.width = dw1, fig.height = dh2}
plot_calibrate(result$biomass_age)
```

\newpage

### Numbers
```{r, fig.width = dw1, fig.height = dh2}
plot_calibrate(result$nums_age)
```


\newpage

### Biomass
```{r, fig.width = dw1, fig.height = dh1}
plot_calibrate(result$biomass)
```

\newpage

## Distribution plots
### Numbers @ age
```{r, fig.width = dw1, fig.height = dh2}
plot_struct(result$nums_age)
```

\newpage

### Biomass @ age
```{r, fig.width = dw1, fig.height = dh2}
plot_struct(result$biomass_age)
```

