---
title: "model-calibration-species"
author: "Alexander Keth"
date: '`r Sys.Date()`'
output: 
  pdf_document:
    toc: true
    number_sections: true
    fig_width: 10
    fig_height: 13
    fig_caption: true
vignette: >
  %\VignetteIndexEntry{model-calibration-species}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

\newpage

NOTE: This vigentte is optimised for longer simulation runs. Therefore the output is not as pleasant
due to the fact that the dummy setas file have a simulation time of 5 years.

In order to use this vignette make sure to render `model-preprocess.Rmd` first. Either save the resulting
list of dataframes as shown in `data-raw/data-vignette-model-preprocess.R` or render both vignettes
`model-preprocess.Rmd` and `model-calibration-species.Rmd` in the same R-instance. Of course,  you can also use 
a personalised version of `mode-preprocess.Rmd`. Please make sure to add all resulting dataframes
to the list of dataframes at the end of the preprocess vignette and change `model-calibration-species.Rmd` accordingly.

```{r}
library("atlantistools")
library("gridExtra")

# You should be able to build the vignette either by clicking on "Knit PDF" in RStudio or with
# rmarkdown::render("model-calibration-species.Rmd")
```

## User Input
This section is used to read in the SETAS dummy files. Please change this accordingly.

```{r}
result <- preprocess
species <- c("Shallow piscivorous fish", "Small planktivorous fish")
```

\newpage

# Overview species 

```{r, echo=FALSE, results = 'asis'}
plots <- lapply(species, plot_species, data_pre = result)
for (i in seq_along(plots)) {
  cat(paste0("## ", species[i]), sep = "\n")
  gridExtra::grid.arrange(plots[[i]])
  cat("\n\n")
}
```









