---
title: "model-calibration"
author: "Alexander Keth"
date: '`r Sys.Date()`'
output: 
  pdf_document:
    toc: true
    number_sections: true
    fig_width: 25
    fig_height: 6
    fig_caption: true
classoption: landscape
geometry: margin = 0.5cm
vignette: >
  %\VignetteIndexEntry{model-calibration}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

\newpage

NOTE: This vigentte is optimised for longer simulation runs. Therefore the output is not as pleasant
due to the fact that the dummy setas file have a running time of 5 years.

In order to use this vignette make sure to render `model-preprocess.Rmd` first. Either save the resulting
list of dataframes as shown in `data-raw/data-vignette-model-preprocess.R` or render both vignettes
`model-preprocess.Rmd` and `model-calibration.Rmd` in the same R-instance. Of course,  you can also use 
a personalised version of `mode-preprocess.Rmd`. Please make sure to add all resulting dataframes
to the list of dataframes at the end of the preprocess vignette and change `model-calibration.Rmd` accordingly.

```{r}
library("atlantistools")
library("ggplot2")
library("gridExtra")

fig_height2 <- 11
gen_labels <- list(x = "Time [years]", y = "Biomass [t]")

# You should be able to build the vignette either by clicking on "Knit PDF" in RStudio or with
# rmarkdown::render("model-calibration.Rmd")
```

## User Input
This section is used to read in the SETAS dummy files. Please change this accordingly.

```{r}
result <- preprocess

d <- system.file("extdata", "setas-model-new-trunk", package = "atlantistools")

# External recruitment data
ex_rec_ssb <- read.csv(file.path(d, "setas-ssb-rec.csv"), stringsAsFactors = FALSE)

# External biomass data
ex_bio <- read.csv(file.path(d, "setas-bench.csv"), stringsAsFactors = FALSE)

# bgm file
bgm <- file.path(d, "VMPA_setas.bgm")
```

\newpage

# Whole system plots!
## Overall biomass
```{r, fig.width = 14, fig.height = 8}
df_bio <- combine_groups(result$biomass, group_col = "species", combine_thresh = 10)
plot <- plot_bar(df_bio)
update_labels(plot, labels = gen_labels)
```

\newpage

## Biomass timeseries
```{r}
plot <- plot_line(result$biomass)
update_labels(plot, labels = gen_labels)
```

\newpage

## Biomass@age timeseries
```{r}
plot <- plot_line(result$biomass_age, col = "agecl")
update_labels(p = plot, labels = c(gen_labels, list(colour = "Ageclass")))
```

\newpage

## Number timeseries
```{r, fig.height = fig_height2}
plot <- plot_line(result$nums)
update_labels(p = plot, labels = list(x = "Time [years]", y = "Numbers"))
```

\newpage

## Number@age timeseries
```{r, fig.height = fig_height2}
plot <- plot_line(result$nums_age, col = "agecl")
update_labels(p = plot, labels = list(x = "Time [years]", y = "Numbers", colour = "Ageclass"))
```

\newpage

## SSB & Recruitment

```{r}
plot_rec(result$ssb_rec, ex_data = ex_rec_ssb)
```

\newpage

## Biomass benchmark
```{r}
names(ex_bio)[names(ex_bio) == "biomass"] <- "atoutput"

data <- result$biomass
data$model <- "atlantis"
comp <- rbind(ex_bio, data, stringsAsFactors = FALSE)

# Show atlantis as first factor!
comp$model <- factor(comp$model, levels = c("atlantis", sort(unique(comp$model))[sort(unique(comp$model)) != "atlantis"]))

# Create plot
plot <- plot_line(comp, col = "model")
update_labels(plot, gen_labels)
```

\newpage

## Biomass benchmark 2
```{r}
plot <- plot_line(result$biomass) %>% update_labels(labels = gen_labels)
plot_add_range(plot, ex_bio)
```

\newpage

## Physics
```{r, fig.width = 30, fig.height = 12}
plot <- plot_line(result$physics, wrap = NULL)
custom_grid(plot, grid_x = "polygon", grid_y = "variable")
```

\newpage

# Physics
```{r, results = 'asis', fig.width = 30, fig.height = 12}
physics <- result$physics %>%
  flip_layers() %>%
  split(., .$variable)

plots <- lapply(physics, plot_line, wrap = NULL) %>% 
  lapply(., custom_grid, grid_x = "polygon", grid_y = "layer")

for (i in seq_along(plots)) {
  cat(paste0("## ", names(plots)[i]), sep = "\n")
  plot <- update_labels(plots[[i]], labels = list(y = names(plots)[i]))
  print(plot)
  cat("\n\n")
}
```

\newpage

## Fluxes 1
```{r, fig.width = 30, fig.height = 12}
plot <- flip_layers(result$flux) %>% 
  plot_line(wrap = NULL, col = "variable")
custom_grid(plot, grid_x = "polygon", grid_y = "layer")
```

\newpage

## Fluxes 2
```{r, fig.width = 30, fig.height = 12}
plot <- flip_layers(result$sink) %>% 
  plot_line(wrap = NULL, col = "variable")
custom_grid(plot, grid_x = "polygon", grid_y = "layer")
```

\newpage

## Relative change of water column height compared to nominal_dz
```{r}
check_dz <- result$dz %>% 
  dplyr::left_join(result$nominal_dz, by = c("polygon", "layer")) %>% 
  dplyr::mutate(check_dz = atoutput.x / atoutput.y) %>% 
  dplyr::filter(!is.na(check_dz)) # remove sediment layer

plot <- plot_line(check_dz, x = "time", y = "check_dz", wrap = "polygon", col = "layer")
update_labels(plot, list(x = "Time [years]", y = expression(dz/nominal_dz)))
```


\newpage

# Calibration plots
## Structural nitrogen
```{r, fig.height = fig_height2}
df_rel <- convert_relative_initial(result$structn_age)
plot <- plot_line(df_rel, col = "agecl")
plot <- update_labels(plot, list(x = "Time [years]", y = expression(SN/SN[init])))
plot_add_box(plot)
```

\newpage

## Reserve nitrogen
```{r, fig.height = fig_height2}
df_rel <- convert_relative_initial(result$resn_age)
plot <- plot_line(df_rel, col = "agecl")
plot <- update_labels(plot, list(x = "Time [years]", y = expression(RN/RN[init])))
plot_add_box(plot)
```

\newpage

## Biomass per ageclass
```{r, fig.height = fig_height2}
df_rel <- convert_relative_initial(result$biomass_age)
plot <- plot_line(df_rel, col = "agecl")
plot <- update_labels(plot, list(x = "Time [years]", y = expression(Biomass/Biomass[init])))
plot_add_box(plot)
```

\newpage

## Eat per ageclass
```{r, fig.height = fig_height2}
df_rel <- convert_relative_initial(result$eat_age)
plot <- plot_line(df_rel, col = "agecl")
plot <- update_labels(plot, list(x = "Time [years]", y = expression(Cons./Cons.[init])))
plot_add_box(plot)
```

\newpage

## Growth per ageclass
```{r, fig.height = fig_height2}
df_rel <- convert_relative_initial(result$growth_age)
plot <- plot_line(df_rel, col = "agecl")
plot <- update_labels(plot, list(x = "Time [years]", y = expression(Growth/Growth[init])))
plot_add_box(plot)
```

\newpage

## Growth in relation to initial conditions
```{r}
plot <- plot_line(result$growth_rel_init, y = "gr_rel", col = "agecl")
update_labels(plot, list(y = expression((Growth - Growth[req])/Growth[req])))
```


\newpage

## Numbers
```{r, fig.height = fig_height2}
df_rel <- convert_relative_initial(result$nums_age)
plot <- plot_line(df_rel, col = "agecl")
plot <- update_labels(plot, list(x = "Time [years]", y = expression(Numbers/Numbers[init])))
plot_add_box(plot)
```

\newpage

## Biomass
```{r}
df_rel <- convert_relative_initial(result$biomass)
plot <- plot_line(df_rel)
plot <- update_labels(plot, list(x = "Time [years]", y = expression(Biomass/Biomass[init])))
plot_add_box(plot)
```

\newpage

# Distribution plots
## Numbers @ age
```{r, fig.height = fig_height2}
df <- agg_perc(result$nums_age, groups = c("time", "species"))
plot <- plot_bar(df, fill = "agecl", wrap = "species")
update_labels(plot, labels = list(x = "Time [years]", y = "Numbers [%]"))
```

\newpage

## Biomass @ age
```{r, fig.height = fig_height2}
df <- agg_perc(result$biomass_age, groups = c("time", "species"))
plot <- plot_bar(df, fill = "agecl", wrap = "species")
update_labels(plot, labels = list(x = "Time [years]", y = "Biomass [%]"))
```

\newpage

# Diet Plots

```{r, echo=FALSE, results = 'asis', fig.width = 14, fig.height = 9}
plots <- plot_diet(result$biomass_consumed, wrap_col = "agecl", combine_thresh = 7)
for (i in seq_along(plots)) {
  cat(paste0("## Diet plot ", i, ": ", names(plots)[i]), sep = "\n")
  gridExtra::grid.arrange(plots[[i]])
  cat("\n\n")
}
```

\newpage

# Spatial Plots 1
```{r, echo=FALSE, results = 'asis', fig.width = 14, fig.height = 9}
plots <- plot_spatial_box(result$biomass_spatial_stanza, bgm_as_df = convert_bgm(bgm = bgm), timesteps = 7)
for (i in seq_along(plots)) {
  cat(paste0("## Spatial Plot ", i, ": ", names(plots)[i]), sep = "\n")
  gridExtra::grid.arrange(plots[[i]])
  cat("\n\n")
}
```

\newpage

# Spatial Plots 2
```{r, echo=FALSE, results = 'asis', fig.width = 14, fig.height = 9}
plots <- plot_spatial_ts(result$biomass_spatial_stanza, bgm_as_df = convert_bgm(bgm = bgm), vol = result$vol)
for (i in seq_along(plots)) {
  cat(paste0("## Spatial Plot ", i, ": ", names(plots)[i]), sep = "\n")
  gridExtra::grid.arrange(plots[[i]])
  cat("\n\n")
}
```

