---
title: "package-demo"
author: "Alexander Keth"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{package-demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library("atlantistools")
library("ggplot2")

d <- system.file("extdata", "setas-model-new-becdev", package = "atlantistools")
```


## Preprocess your atlantis outoput
The first step with `atlantistools` is generally the preprocessing of the output of your atlantis simulation. 
During this process an extensive set of dataframes is created from the general-
and production netcdf files (e.g. individual weight, biomass, numbers at age ...). These dataframes
can be directly passed to various plotting routines.

```{r, eval = FALSE}
setas <-   preprocess(dir = d,
                      nc_gen = "outputSETAS.nc",
                      nc_prod = "outputSETASPROD.nc",
                      dietcheck = "outputSETASDietCheck.txt",
                      yoy = "outputSETASYOY.txt",
                      ssb = "outputSETASSSB.txt",
                      prm_biol = "VMPA_setas_biol_fishing_New.prm",
                      prm_run = "VMPA_setas_run_fishing_F_New.prm",
                      bps = load_bps(dir = d, fgs = "SETasGroups.csv", 
                                     init = "init_vmpa_setas_25032013.nc"),
                      fgs = "SETasGroups.csv",
                      select_groups = c("Planktiv_S_Fish", "Pisciv_S_Fish", 
                                        "Cephalopod", "Diatom", "Zoo"),
                      bboxes = get_boundary(boxinfo = load_box(dir = d, 
                                                               bgm = "VMPA_setas.bgm")),
                      check_acronyms = TRUE,
                      modelstart = "1991-01-01",
                      out = "preprocess.Rda",
                      report = FALSE,
                      save_to_disc = FALSE)
```

```{r, echo = FALSE}
setas <- suppressMessages(
  preprocess(dir = d,
             nc_gen = "outputSETAS.nc",
             nc_prod = "outputSETASPROD.nc",
             dietcheck = "outputSETASDietCheck.txt",
             yoy = "outputSETASYOY.txt",
             ssb = "outputSETASSSB.txt",
             prm_biol = "VMPA_setas_biol_fishing_New.prm",
             prm_run = "VMPA_setas_run_fishing_F_New.prm",
             bps = load_bps(dir = d, fgs = "SETasGroups.csv", 
                            init = "init_vmpa_setas_25032013.nc"),
             fgs = "SETasGroups.csv",
             select_groups = c("Planktiv_S_Fish", "Pisciv_S_Fish", 
                               "Cephalopod", "Diatom", "Zoo"),
             bboxes = get_boundary(boxinfo = load_box(dir = d, 
                                                      bgm = "VMPA_setas.bgm")),
             check_acronyms = TRUE,
             modelstart = "1991-01-01",
             out = "preprocess.Rda",
             report = FALSE,
             save_to_disc = FALSE))
```

You can also directly pass the boundary boxes or epibenthic groups as vectors if you do not intend to use the build in
functions `load_bps` or `get_boundary`. However, I highly recommend to use these build in functions
as they ensure the correct output independant of model formulation.

```{r, eval = FALSE}
setas <- preprocess(dir = d,
                    nc_gen = "outputSETAS.nc",
                    nc_prod = "outputSETASPROD.nc",
                    dietcheck = "outputSETASDietCheck.txt",
                    yoy = "outputSETASYOY.txt",
                    ssb = "outputSETASSSB.txt",
                    prm_biol = "VMPA_setas_biol_fishing_New.prm",
                    prm_run = "VMPA_setas_run_fishing_F_New.prm",
                    bps = c("Filter_Shallow", 
                            "Filter_Other", 
                            "Filter_Deep", 
                            "Benthic_grazer", 
                            "Macrobenth_Deep", 
                            "Megazoobenthos", 
                            "Macrobenth_Shallow", 
                            "Macroalgae"),
                    fgs = "SETasGroups.csv",
                    select_groups = c("Planktiv_S_Fish", "Pisciv_S_Fish", 
                                      "Cephalopod", "Diatom", "Zoo"),
                    bboxes = c(0, 6, 7, 8, 9, 10),
                    check_acronyms = TRUE,
                    modelstart = "1991-01-01",
                    out = "preprocess.Rda",
                    report = FALSE,
                    save_to_disc = FALSE)
```

The data can be stored on your machine if `save_to_disc` is set
to `TRUE`. The data is stored as `*.rda` format as a list of dataframes which all share the same format.
This is done to save storage space and to make sure that the data structure is not altered when it is imported into R.
If you want to work with external software in further steps (E.g. Excel) you should convert the dataframes to `*.txt` or `*.csv` file.
In addition you should always set `report` to `TRUE` to activate the progression bars during
data import and transformation since this process may take some time for complex models (many boxes/groups)
or long simulations.

Most functions in `atlantistools` expect a directory to search for specific model files. There are 2 ways to pass the directory:

All your model files are stored in one directory. The files can now be accesed via their name. The directory can be set/passed in two ways:

Set your working directory in R to your atlantis model folder. In this case you don't have to pass the directory.
Pass the directory of your Atlantis folder as Object (this is the approach which is used here) to the argument `dir`.
  
In case you are using different folders for generic model files and output files the directory can be set as described above. However you need to
implement the output subfolder into the name of your outputfiles:

```{r, eval = FALSE}

setas <- preprocess(dir = d,
                    nc_gen = "output/outputSETAS.nc",
                    nc_prod = "output/outputSETASPROD.nc",
                    dietcheck = "output/outputSETASDietCheck.txt",
                    yoy = "output/outputSETASYOY.txt",
                    ssb = "output/outputSETASSSB.txt",
                    prm_biol = "VMPA_setas_biol_fishing_New.prm",
                    prm_run = "VMPA_setas_run_fishing_F_New.prm",
                    bps = load_bps(dir = d, fgs = "SETasGroups.csv", 
                                   init = "init_vmpa_setas_25032013.nc"),
                    fgs = "SETasGroups.csv",
                    select_groups = c("Planktiv_S_Fish", "Pisciv_S_Fish", 
                                      "Cephalopod", "Diatom", "Zoo"),
                    bboxes = get_boundary(boxinfo = load_box(dir = d, 
                                                             bgm = "VMPA_setas.bgm")),
                    check_acronyms = TRUE,
                    modelstart = "1991-01-01",
                    out = "preprocess.Rda",
                    report = FALSE,
                    save_to_disc = FALSE)
```

## Create a map of your model for Bec!

Always add a spatial representation of your model if you have issues with model tuning. This makes the life of Bec and Beth much easier.

```{r, fig.width = 7, fig.height = 5}
bgm_data <- convert_bgm(dir = d, bgm = "VMPA_setas.bgm")
plot_boxes(bgm_data)
```

## Plot physics and fluxes

```{r, fig.width = 7, fig.height = 10}
plot_physics(setas$physics)
```

```{r, fig.width = 7, fig.height = 5}
plot_flux(setas$flux)
```


## Plot biomass timeseries (or any other time series!)

```{r, fig.width = 10, fig.height = 3}
plot_ts(setas$biomass)
```

## Create reference plot for individual weight and biomass

```{r, fig.width = 7, fig.height = 5}
plot_calibrate(data = setas$structn_age)
```

The function can also be used with non-age structured data.

```{r, fig.width = 10, fig.height = 3}
plot_calibrate(data = setas$biomass)
```

## Visualise cohort distributions for numbers and biomass over time

```{r, fig.width = 7, fig.height = 5}
plot_struct(data = setas$biomass_age)
plot_struct(data = setas$nums_age)
```

## Plot DietCheck.txt

Please note the plots will look much nicer if the simulation period is elongated.
I only ran the SETAS model provided with `atlantistools` for 3 years to minimise the size of the package.

```{r, fig.show = 'hold', fig.width = 7, fig.height = 5}
# Create a plot for each predator in your model.
feeding_plots <- plot_dietcheck(data = setas$diet, combine_thresh = 0.03)
feeding_plots[[1]]
```

```{r, eval = FALSE}
# Save all plots to disc in multiple pdfs!
for (i in seq_along(feeding_plots)) {
  ggsave(filename = paste0("feeding", i, ".pdf"), 
         plot = feeding_plots[[i]], path = d, width = 14, height = 10)
}

# Save all plots to disc in one pdf!
pdf(file.path(d, "feeding.pdf"), width = 14, height = 10)
for (i in seq_along(feeding_plots)) {
  print(feeding_plots[[i]])
}
dev.off()
```

## Compare recruitment with recruitment data

Please note that recruitment in Atlantis differs from recruitment in most models due to changes of recruitment
ages (which is generally age 0-3 in stock assessment models). Thanks @ Isaac for pointing that out.

As the Setas model comes without recruitment data some noise is applied to the output files to demonstrate the plots.

```{r, fig.width = 7, fig.height = 5}
ex_data <- setas$ssb_rec
ex_data$species <- as.character(ex_data$species)
ex_data$rec <- ex_data$rec * runif(n = nrow(ex_data), min = 0.8, max = 1.2)
ex_data$ssb <- ex_data$ssb * runif(n = nrow(ex_data), min = 0.8, max = 1.2)
ex_data$model <- "test_model"
plot_rec(data = setas$ssb_rec, ex_data)
```

## Compare your model run with other models/data!

Again, noise is added due to a lack of external data.

```{r, fig.width = 10, fig.height = 4}
data_comp <- setas$biomass
data_comp$biomass <- data_comp$atoutput * runif(n = nrow(data_comp), min = 0.8, max = 1.2)
data_comp$atoutput <- NULL
data_comp$model <- "test_model"
plot_bench(data = setas$biomass, ex_data = data_comp)
```

## Change parameter values!

Let's say you want to change a parameter in your biological parameter file. For example the recruit 
weights for a specific fish. In our case `Small planktivorous fish` (Code = `FPS`). You can either
change the existing valu by multiplication with a factor or set a new absolut value. In our case
we increase the existing value by a factor of 2.

```{r}
new_prm <-  change_prm(dir = d, 
                       prm_biol = "VMPA_setas_biol_fishing_New.prm", 
                       select_acronyms = "FPS", 
                       roc = 2, 
                       parameter = "KWRR", 
                       save_to_disc = FALSE)
```

You can use `extract_prm` to check the format of the resulting *.prm. However, I checked the function
thoroughly so you would normally set `save_to_disc` to `TRUE` to overwrite your existing
parameter file with the newly generated one Please make sure to backup your original files beforehand.

```{r}
prm_old <- readLines(con = file.path(d, "VMPA_setas_biol_fishing_New.prm"))
extract_prm(chars = prm_old, variable = "KWRR_FPS")
extract_prm(chars = new_prm, variable = "KWRR_FPS")
```

You can also pass a vector of groups to change the parameter for multiple functional groups with a single call to `change_prm`.

```{r}
new_prm <-  change_prm(dir = d, 
                       prm_biol = "VMPA_setas_biol_fishing_New.prm", 
                       select_acronyms = c("FPL", "FPO", "FPS", "FVD", "FVV", "FVS", "FVB", "FVT", "FVO"), 
                       roc = runif(n = 9, min = 2, max = 5), 
                       parameter = "KWRR", 
                       save_to_disc = FALSE)
```

This helps to automate the tuning and calibration of your model. In addition it is a useful tool to make the tuning
reproduceable and transparent for other modelers.

In case you want to change the parameter values for cohort structured groups with multiple values per parameter
(E.g. Clearance rate and mum for fish groups) please use `change_prm_cohrt`. Please note that the function only
works with parameters whose values are stored in the next row following the flag in the *.prm file.

```{r}
new_prm <-  change_prm_cohort(dir = d, 
                              prm_biol = "VMPA_setas_biol_fishing_New.prm", 
                              select_acronyms = c("FPL", "FPO"), 
                              roc = matrix(rep(2, times = 20), nrow = 2, ncol = 10), 
                              parameter = "C", 
                              save_to_disc = FALSE)
```


